CUTIE: Learning to Understand Documents with Convolutional Universal Text Information Extractor

### 2019-04-04 Article/ms.tex
CUTIE paper final version

%%%%%%%%% ABSTRACT
\begin{abstract}
   Extracting key information from documents, such as receipts or invoices, and preserving the interested texts to structured data is crucial in the document-intensive streamline processes of office automation in areas that includes but not limited to accounting, financial, and taxation areas. To avoid designing expert rules for each specific type of document, some published works attempt to tackle the problem by learning a model to explore the semantic context in text sequences based on the Named Entity Recognition (NER) method in the NLP field. In this paper, we propose to harness the effective information from both semantic meaning and spatial distribution of texts in documents. Specifically, our proposed model, Convolutional Universal Text Information Extractor (CUTIE), applies convolutional neural networks on gridded texts where texts are embedded as features with semantical connotations. We further explore the effect of employing different structures of convolutional neural network and propose a fast and portable structure. We demonstrate the effectiveness of the proposed method on a dataset with up to $6,980$ labelled receipts, without any pre-training or post-processing, achieving state of the art performance that is much higher than BERT but with only $1/10$ parameters and without requiring the $3,300$M word dataset for pre-training. Experimental results also demonstrate that the CUTIE being able to achieve state of the art performance with much smaller amount of training data.
\end{abstract}

@@

Both CUTIE-A and CUITE-B conducts semantical meaning encoding process with a word embedding layer in the very beginning stage. The cross entropy loss function is applied to compare the predicted token class grid and the ground truth grid.

@@

\begin{table}
	\caption{Statistic of the numbers of labelled receipt document images and key information classes in the dataset.}
\begin{center}
\begin{tabular}{l | c | c | c}
	 & Training Set & Test Set & \#classes \\
	\hline
	ME & 1109 & 475 & 9 \\
	Taxi & 1125 & 375 & 6 \\
	Hotel & 1125 & 375 & 9 \\
\end{tabular}
\end{center}
	\label{tab:dataset}
\end{table}

\begin{table*}
	\caption{Performance comparison on different types of documents. (AP/softAP)}
\begin{center}
\begin{tabular}{l | c | c | c | c}
	Method & \#Params & Taxi & ME & Hotel \\
	\hline
	CloudScan\cite{cloudscan} & - & 82 / - & 64 / - & 60 / - \\
	BERT\cite{bert} & 110M & 88.1 / - & 80.1 / - & 71.7 / - \\
	CUTIE-A & 67M & \\
	CUTIE-B & 14M & 93.3 / 96.9 & 81.0 / 88.9 & 75.8 / 87.1 \\
	%CUTIE-A split & 67M & 92.9 / 96.7 & 81.4 / 91.6 & 72.3 / 95.0 \\
	%CUTIE-B split & 14M & 95.2 / 97.6 & 84.3 / 92.4 & 70.8 / 89.4 \\
\end{tabular}
\end{center}
	\label{tab:comparison}
\end{table*}

@@

We compare the performance of the proposed method with two state of the art methods CloudScan \cite{cloudscan} and BERT \cite{bert}. For comparison, the Cloud Scan model for SROIE is trained from scratch but with several expert designed features as described in \cite{cloudscan}. The BERT model for SROIE is transform learned with the Google released base model that is pre-trained on a huge dataset with $3,300$M words \cite{bert,bertgit}. It is worth noting that the CloudScan based method and BERT based method were implemented by other teams. For that reason, there is minor difference in numbers of training / test data. To provide a fair comparison, we also provide ablation study to evaluate the influence of number of training images in CUTIE. We will also provide results generated using the same dataset in the next version of this paper.

We use a learning rate of $1e-3$ with Adam optimizer and step decay learning strategy. The learning rate is dropped to $1e-4$ and $1e-5$ on the $15,000$-th and $30,000$-th steps, respectively. The training is terminated within $40,000$ steps with batch size of $32$. We train our model on Tesla V100 GPU where $11$ to $19$GB memories is used depending the configuration of the model framework and size of the dataset. Instance normalization is involved to facilate training. Our model is trained end-to-end without piecewise pretraining of any component. The default embedding size is $128$, target augmentation shape is $64$ for both row and column. The dataset is split as training set and test set with ratio of $75:25$. No pre-processing or post-processing are involved in CUTIE.

@@

\subsection{Dataset}
The dataset contains $6,980$ annotated scanned Spanish receipt documents, including taxi receipts, meals entertainment (ME) receipts, and hotel receipts, with $9$ different key information classes, as reported in Table \ref{tab:dataset}. We generate the texts and corresponding bounding boxes with Google's OCR API. Each text and their bounding box is manually labelled as one of the $9$ different classes: 'DontCare', 'VendorName', 'VendorTaxID', 'InvoiceDate', 'InvoiceNumber', 'ExpenseAmount', 'BaseAmount', 'TaxAmount', and 'TaxRate'. We then employ the tokenizer introduced in Section \ref{pm} to segment texts into minimum token units, where text bounding boxes are also segmented accordingly.

@@

\subsection{Overall Performance}
We report results of our method in terms of AP and compare with other state of the art methods in Table \ref{tab:comparison}. We also provides softAP results for CUTIE-A and CUTIE-B in Table \ref{tab:comparison}, where both the softAP of CUTIE-A and CUTIE-B exceeds their AP by a large margin.  Examples of inference results are illustrated in Figure \ref{fig:result}. Our big network CUTIE-A achieves $91.2\%$ AP and $98.1\%$ softAP on taxi receipts, $84.8\%$ AP and $93.5\%$ softAP on meals entertainment receipts, and $62.7\%$ AP and $88.0\%$ softAP on hotel receipts, outperforming other methods in almost all of the test cases. Compared to CloudScan, our big network CUTIE-A improves AP by $20.8\%$ AP on meals entertainment receipts, $9.2\%$ on taxi receipt, and $2.7\%$ on hotel receipts; our small network CUTIE-B improves AP by $13.2$ on taxi receipts, $20.3\%$ points on meals entertainment receipts, and $10.8\%$ on hotel receipts. Furthermore, compared to BERT, where model is pre-trained on a dataset with $3,300M$ words and transfer learned on our dataset, our big network CUTIE-B improves AP by $4.2\%$ and $4.5\%$ in taxi receipts and meals entertainment while using only $1/2$ parameters, our small network CUTIE-B improves AP by $8.2\%$ and $4.0\%$ in taxi receipts and meals entertainment but with much less complexity and smaller model size with only $1/9$ parameters without requiring a huge dataset for model pretraining. We will further prove in Section \ref{parameters} that CUTIE-B is also able to achieve state of the art performance with only $1/10$ of BERT's parameters. Although CUTIE-B is smaller in capacity, it outperforms CUTIE-A in several assessment criterias. This is because CUTIE-B enlarges field of view by employing the atrous convolution rather than the pooling or striding processes, the CUTIE-B model has larger field of view and better understanding of tokens' relative spatial relationship since no restrictions is applied on the feature maps shapes.

It is worth noting that the difference of AP in the hotel receipt column actually leads to interesting findings. One of the finding is that if we neglect the 'VendorName' class in the evaluation criteria, the AP is increased to from $70.8\%$ to $76.9\%$ and the softAP drops to $83.8\%$, indicating that CUTIE is capable of extracting the interested 'VendorName' texts but sometimes involves texts that are not in the ground truth label. Another finding is that the hotel receipts are quite different from the meals entertainment receipts and the taxi receipts, where key information appears multi times in different areas of the receipt, whereas the human labelers tend to label only one of their appearances. We look deeper into this in the following part of this section by analyzing some inference result cases.

Typical examples of receipts with low AP but high softAP score are shown in Figure \ref{fig:falsepositive}. Most of the false positive cases occur in the 'VenderName' class, where names tend to greatly vary and leads to difficulty in model inference. However, it is not hard to find that these false positives can be easily avoided by appending a dictionary based post processor to the key information extractor. One rare false positive case, the third receipt in the first row, is a 'L' letter being mis-recognized as '1' by the employed OCR engine. Although the letter shows distinct appearance from the other digits in the scanned receipt image, it is mis-interpreted as one part of the 'BaseAmount' class due to its close spatial location to the digits and it-self being a digit. It also can be seen in the first receipt in the second row that several spatial distant digits were wrongly predicted as in the 'BaseAmount' class. Although these are rare cases in our test set, it still suggests that incorporating image-level information may further boost the inference accuracy inspite of the already involved semantical and spatial features.

@@

\subsection{Ablation Studies}
Although we have demonstrated extremely strong empirical results, the results presented are achieved by combination of each aspect of the CUTIE framework. In this section, we perform ablation studies over a number of facets of CUTIE in order to have a better understanding of their relative importance. CUTIE-B is employed as the default model with grid augmentation, embedding size of $128$, and $75\%$ of dataset as training data.

@@

\subsubsection{Impact of Semantical Information Capacity}
\label{parameters}
Next, we evaluate the impact of semantical information of CUTIE by comparing evaluation with different word embedding sizes. As reported in Table \ref{tab:embedding}, CUTIE performs in a bell shape curve as the embedding size increases. The best performance is achieved by CUTIE-B with $85.0\%$ in AP and $92.9\%$ in softAP using embedding size of $64$. One interesting finding is that the CUTIE model achieves good performance even with limited semantical information capacity. We infer the reason is that, for the SROIE problem with 9 key information classes, there are a small amount of key tokens provide the majority of contribution to the model inference and the model can achieve good performance by paying special attention to these key tokens. It is also indicated in the result that too large embedding size also decreases the model performace. To further enhance the model performance, it might help to randomly mask out certain texts during the grid positional mapping process, such that the model learns inter-text relationship better with / without certain words present in the grid and further improves in generality.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
   Extracting key information from documents, such as receipts or invoices, and preserving the interested texts to structured data is crucial in the document-intensive streamline processes of office automation in areas that includes but not limited to accounting, financial, and taxation areas. To avoid designing expert rules for each specific type of document, some published works attempt to tackle the problem by learning a model to explore the semantic context in text sequences based on the Named Entity Recognition (NER) method in the NLP field. In this paper, we propose to harness the effective information from both semantic meaning and spatial distribution of texts in documents. Specifically, our proposed model, Convolutional Universal Text Information Extractor (CUTIE), applies convolutional neural networks on gridded texts where texts are embedded as features with semantical connotations. We further explore the effect of employing different structures of convolutional neural network and propose a fast and portable structure. We demonstrate the effectiveness of the proposed method on a dataset with up to $4,484$ labelled receipts, without any pre-training or post-processing, achieving state of the art performance that is much higher than BERT but with only $1/10$ parameters and without requiring the $3,300$M word dataset for pre-training. Experimental results also demonstrate that the CUTIE being able to achieve state of the art performance with much smaller amount of training data.
\end{abstract}

@@

Both CUTIE-A and CUITE-B conducts semantical meaning encoding process with a word embedding layer in the very beginning stage. Dropout is applied on the embedding layer to enhance the CUTIE's generalization ability. The cross entropy loss function is applied to compare the predicted token class grid and the ground truth grid.

@@

\begin{table}
	\caption{Statistic of the numbers of labelled receipt document images and key information classes in the dataset.}
\begin{center}
\begin{tabular}{l | c | c | c}
	 & Training Set & Test Set & \#classes \\
	\hline
	ME & 1109 & 375 & 9 \\
	Taxi & 1125 & 375 & 6 \\
	Hotel & 1125 & 375 & 9 \\
\end{tabular}
\end{center}
	\label{tab:dataset}
\end{table}

\begin{table*}
	\caption{Performance comparison on different types of documents. (AP/softAP)}
\begin{center}
\begin{tabular}{l | c | c | c | c}
	Method & \#Params & Taxi & ME & Hotel \\
	\hline
	CloudScan\cite{cloudscan} & - & 82 / - & 64 / - & 60 / - \\
	BERT\cite{bert} & 110M & 88.1 / - & 80.1 / - & 71.7 / - \\
	CUTIE-A & 67M & 90.8 / 97.2 & 77.7 / \textbf{91.4} & 69.5 / \textbf{87.8} \\
	CUTIE-B & 14M & \textbf{94.0} / \textbf{97.3} & \textbf{81.5} / 89.7 & \textbf{74.6} / 87.0 \\
	%CUTIE-A split & 67M & 92.9 / 96.7 & 81.4 / 91.6 & 72.3 / 95.0 \\
	%CUTIE-B split & 14M & 95.2 / 97.6 & 84.3 / 92.4 & 70.8 / 89.4 \\
\end{tabular}
\end{center}
	\label{tab:comparison}
\end{table*}

@@

We compare the performance of the proposed method with two state of the art methods CloudScan \cite{cloudscan} and BERT \cite{bert}. For comparison, the Cloud Scan model for SROIE is trained from scratch but with several expert designed features as described in \cite{cloudscan}. The BERT model for SROIE is transform learned with the Google released base model that is pre-trained on a huge dataset with $3,300$M words \cite{bert,bertgit}. To provide a fair comparison, $4,484$ samples with around $1,500$ each from Taxi, Meals Entertainment (ME), and Hotel, where $3,375$ samples are used for training and $1,125$ are used for testing, as reported in Table \ref{tab:dataset}. A single model is trained on the dataset for all these $3$ types of document receipts either for CloudScan, BERT, or CUTIE.

We use a learning rate of $1e-3$ with Adam optimizer and step decay learning strategy. The learning rate is dropped to $1e-4$ and $1e-5$ on the $15,000$-th and $30,000$-th steps, respectively. The training is terminated within $40,000$ steps with batch size of $32$. We train our model on Tesla V100 GPU where $11$ to $19$GB memories is used depending the configuration of the model framework and size of the dataset. Instance normalization and hard negative mining are involved to facilate training. Dropout is applied with keep probability of $0.9$. Our model is trained end-to-end without piecewise pretraining of any component. The default embedding size is $128$, target augmentation shape is $64$ for both row and column. The dataset is split as training set and test set with ratio of $75:25$. No pre-processing or post-processing are involved in CUTIE.

@@

\subsection{Dataset}
The dataset contains $4,484$ annotated scanned Spanish receipt documents, including taxi receipts, meals entertainment (ME) receipts, and hotel receipts, with $9$ different key information classes. We generate the texts and corresponding bounding boxes with Google's OCR API. Each text and their bounding box is manually labelled as one of the $9$ different classes: 'DontCare', 'VendorName', 'VendorTaxID', 'InvoiceDate', 'InvoiceNumber', 'ExpenseAmount', 'BaseAmount', 'TaxAmount', and 'TaxRate'. We then employ the tokenizer introduced in Section \ref{pm} to segment texts into minimum token units, where text bounding boxes are also segmented accordingly.

@@

\subsection{Overall Performance}
We report results of our method in terms of AP and compare with other state of the art methods in Table \ref{tab:comparison}. We also provides softAP results for CUTIE-A and CUTIE-B in Table \ref{tab:comparison}, where both the softAP of CUTIE-A and CUTIE-B exceeds their AP by a large margin.  Examples of inference results are illustrated in Figure \ref{fig:result}. Our big network CUTIE-A achieves $90.8\%$ AP and $97.2\%$ softAP on taxi receipts, $77.7\%$ AP and $91.4\%$ softAP on meals entertainment receipts, and $69.5\%$ AP and $87.8\%$ softAP on hotel receipts. Compared to CloudScan, CUTIE-A and CUTIE-B outperforms in every single test case. Furthermore, compared to BERT, where model is pre-trained on a dataset with $3,300$M words and transfer learned on our dataset, our big network CUTIE-A improves AP by $2.7\%$ AP on taxi receipt but less accurate on other type of documents while using only $1/2$ parameters; our small network CUTIE-B improves AP by $5.9\%$ on taxi receipts, $1.4\%$ points on meals entertainment receipts, and $2.9\%$ on hotel receipts, outperforming other methods in all of the test cases but with much less complexity and smaller model size with only $1/9$ parameters and without requiring a huge dataset for model pretraining. We will further prove in Section \ref{parameters} that CUTIE-B is also able to achieve state of the art performance with only $1/10$ of BERT's parameters. Although CUTIE-B is smaller in capacity, it outperforms CUTIE-A in several assessment criterias. This is because CUTIE-B enlarges field of view by employing the atrous convolution rather than the pooling or striding processes, the CUTIE-B model has larger field of view and better understanding of tokens' relative spatial relationship since no restrictions is applied on the feature maps shapes.

It is worth noting that the difference of AP and softAP leads to interesting findings. One of the finding is that CUTIE is capable of extracting the interested texts but sometimes involves texts that are not in the ground truth. Another finding is that the hotel receipts are quite different from the meals entertainment receipts and the taxi receipts, where key information appears multi times in different areas of the receipt, whereas the human labelers tend to label only one of their appearances. We look deeper into this in the following part of this section by analyzing some inference result cases.

Typical examples of receipts with low AP but high softAP score are shown in Figure \ref{fig:falsepositive}. Most of the false positive cases occur in the 'VenderName' class, where names tend to vary greatly and leads to difficulty in model inference. However, it is not hard to find that these false positives can be easily avoided by appending a dictionary based post processor to the key information extractor. One rare false positive case, the third receipt in the first row, is a 'L' letter being mis-recognized as '1' by the employed OCR engine. Although the letter shows distinct appearance from the other digits in the scanned receipt image, it is mis-interpreted as one part of the 'BaseAmount' class due to its close spatial location to the digits and it-self being a digit. It also can be seen in the first receipt in the second row that several spatial distant digits were wrongly predicted as in the 'BaseAmount' class. Although these are rare cases in our test set, it still suggests that incorporating image-level information may further boost the inference accuracy inspite of the already involved semantical and spatial features.

@@

\subsection{Ablation Studies}
Although we have demonstrated extremely strong empirical results, the results presented are achieved by combination of each aspect of the CUTIE framework. In this section, we perform ablation studies over a number of facets of CUTIE in order to have a better understanding of their relative importance. CUTIE-B is employed as the default model with grid augmentation, embedding size of $128$. In this section, we use all data of ME, where $75\%$ of them are used train the model and the rest $25\%$ for test.

@@

\subsubsection{Impact of Semantical Information Capacity}
\label{parameters}
Next, we evaluate the impact of semantical information of CUTIE by comparing evaluation with different word embedding sizes. As reported in Table \ref{tab:embedding}, CUTIE performs in a bell shape curve as the embedding size increases. The best performance is achieved by CUTIE-B with $85.0\%$ in AP and $92.9\%$ in softAP using embedding size of $64$. One interesting finding is that the CUTIE model achieves good performance even with limited semantical information capacity. We infer the reason is that, for the SROIE problem with 9 key information classes, there are a small amount of key tokens provide the majority of contribution to the model inference and the model can achieve good performance by paying special attention to these key tokens. It is also indicated in the result that too large embedding size also decreases the model performace.
